---
title: "Mastodon network analysis"
author: "Gábor T. Mozol"
date: "June, 2024"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Mastodon.Social analysis

This notebook is responsible for the comparative analysis of the mastodon.social data set acquired through scraping the network with the library rtoot and a public user application's API key. Unlike the Truth Social data set, the mastodon.social data set is read in and analysed in situ.

The mastodon.social graph is compared with model graphs:
* Erdős-Rényi or Gnp graph.
* Watts-Strogatz, a graph model with small-world properties.
* Barabási-Albert, a preferential attachment graph model.

The graph models are a 1:1 scale version of the estimated total size of the mastodon.social network size in the Autumn of 2022 (note that the data gathered is up to date as of June 2024). These models are then scraped using a similar BFS method as by Gerard et al. (2023). Based on our assumptions these samples are idealized reproductions of the Truth Social sample.

## Libraries
The notebook uses the following libraries.

**Run**
```{r warning=FALSE}
library(ggraph)
library(igraph)
library(dplyr)
```

## Project setup
First, the wd needs to be set up.
Make sure to set you wd to the project folder.

**Read before run**: uncomment create_project and specify new Rproj file if this is a new setup. Otherwise leave it be.
```{r}
my_wd <- getwd()
project <- file.path(my_wd)
#usethis::create_project(project, open = FALSE)
usethis::proj_set(project)
source(file.path(my_wd,"functions/functions_analysis.R"))
```

## The Mastodon sample
The mastodon.social sample is stored in to .RDS files as the result of the scraping. **nodes.RDS** contains the nodes, while **edges.RDS** contains the edges of the graph.
There are a few duplicates (below 5%) in the tables that need to be cleaned up first.

Then the data will be read into a graph and analysis function will return its descriptive statistics.

### Read and clean the data

First, read in the data already gathered. (see mastodon_scrape.Rmd for details)

**Run**
```{r}
nodes <- readRDS(file.path(my_wd, "data/mastodon_data/nodes.rds"))
edges <- readRDS(file.path(my_wd, "data/mastodon_data/edges.rds"))
```

Second, remove duplicates.

**Run**
```{r}
unique_edges <- edges %>%
  distinct(follower, followee)
unique_nodes <- nodes %>%
  distinct(id, .keep_all = TRUE)
```

Third, check if all edge ids are in nodelist, append if not.

**Run**
```{r}
ids_to_add <- union(setdiff(edges$follower, nodes$id), setdiff(edges$followee, nodes$id))

additional_nodes <- data.frame(
  id = ids_to_add,
  name = NA_character_,
  followers_count = NA_integer_
)

unique_nodes <- bind_rows(unique_nodes, additional_nodes)
```

### Create graph representation

**Read before run**: As a version of the graph is stored separately in a .csv, only run this, if it is outdated. (check next chunk)
```{r}
g_ma <- graph_from_data_frame(d = unique_edges, vertices = unique_nodes, directed = TRUE)
```

**Read before run**: If you want to use the previously generated graph, you can load it in this way.
```{r}
g_ma <- read_graph(file.path(my_wd,"/graph/ma.csv"), "edgelist", directed=TRUE)
```

### Analyze

**Run**
```{r}
g_ma_analysis <- graph_main_analysis(g_ma)
```

**Run**
```{r}
cat("Order: ", g_ma_analysis$order, "\n")
cat("Size: ", g_ma_analysis$size, "\n")
cat("Degree range: ", g_ma_analysis$degree_min, " - ", g_ma_analysis$degree_max, "\n")
cat("Indegree range: ", g_ma_analysis$indegree_min, " - ", g_ma_analysis$indegree_max, "\n")
cat("Outdegree range: ", g_ma_analysis$outdegree_min, " - ", g_ma_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", g_ma_analysis$clusters$weak, " / ", g_ma_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", g_ma_analysis$e_con, "\n")
cat("Vertex connectivity: ", g_ma_analysis$v_con, "\n")
cat("Diameter: ", g_ma_analysis$diameter, "\n")
cat("Average path length: ", g_ma_analysis$apl, "\n")
cat("Transitivity: ", g_ma_analysis$transitivity, "\n")
```

### Save
**Read before run**: Watch out, don't unnecessarily overwrite the existing .csv graph representation.
```{r}
write_graph(g_ma,file.path(my_wd,"/graph/ma.csv"),"edgelist")
```

### Write out Mastodon information for later analysis

**Run**
```{r}
ma_sample_order <- g_ma_analysis$order
ma_sample_size <- g_ma_analysis$size
ma_sample_p <- 2*ma_sample_size/(ma_sample_order*(ma_sample_order-1))
scale_sample_pop <- g_ma_analysis$order/950000
ma_pop_order <- round(ma_sample_order/scale_sample_pop,0)
ma_pop_size <- round(ma_sample_size/scale_sample_pop,0)
```

## Model graphs

**Run**
```{r}
scale <- 1/2
```

### Erdős-Rényi or Gnp model graph analysis

#### Generate
The parameters required are n and p. The value of n is the order of the network, while p is the probability of a connection between any pair of nodes. This probability was calculated under the assumption that an average TS user might be connected to a 100 other users as follower or followee on average. We have the p value of the Truth Social network, ideally the generated network has a similar p value (i.e. within an order of magnitude).

**Read before run**: This chunk generates a new model graph, but there is a stored one. Only use this if you wish to overwrite later. (check next chunk)
```{r}
n <- round(ma_pop_order * scale,0)
p <- 0.0000526 #(average 50 follower relationships in a 950.000 network)
set.seed(n)
g_pop_er <- sample_gnp(n, p, directed = TRUE)

g_sample_er <- bfs_subgraph(g_pop_er, scale = scale_sample_pop)
er_sample_order <- vcount(g_sample_er)
er_sample_size <- ecount(g_sample_er)
er_sample_p <- 2*er_sample_size/(er_sample_order*(er_sample_order-1))
cat("Sample/pop scale: ", vcount(g_sample_er)/vcount(g_pop_er),"\n")
cat("P value: ", er_sample_p,"\n")
cat("Target p value: ", ma_sample_p,"\n")
```

**Read before run**: If you want to use the previously generated graph, you can load it in this way.
```{r}
g_sample_er <- read_graph(file.path(my_wd,"/model_graphs/gnp_ma.csv"), "edgelist", directed=TRUE)
```

#### Analyze
The analysis function finds the information on the model network.

**Run**
```{r}
er_sample_analysis <- graph_main_analysis(g_sample_er)
```
**Run**
```{r}
cat("Order: ", er_sample_analysis$order, "\n")
cat("Size: ", er_sample_analysis$size, "\n")
cat("Degree range: ", er_sample_analysis$degree_min, " - ", er_sample_analysis$degree_max, "\n")
cat("Indegree range: ", er_sample_analysis$indegree_min, " - ", er_sample_analysis$indegree_max, "\n")
cat("Outdegree range: ", er_sample_analysis$outdegree_min, " - ", er_sample_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", er_sample_analysis$clusters$weak, " / ", er_sample_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", er_sample_analysis$e_con, "\n")
cat("Vertex connectivity: ", er_sample_analysis$v_con, "\n")
cat("Diameter: ", er_sample_analysis$diameter, "\n")
cat("Average path length: ", er_sample_analysis$apl, "\n")
cat("Transitivity: ", er_sample_analysis$transitivity, "\n")
```

#### Save
**Read before run**: Watch out, don't unnecessarily overwrite the existing .csv graph representation.
```{r}
write_graph(g_sample_er,file.path(my_wd,"/model_graphs/gnp_ma.csv"),"edgelist")
```

### Watts-Strogatz or Small World Graph analysis
Disclaimer: The Watts-Strogatz network is NOT directed. So in and outdegree values cannot be compared by default.

#### Generate
The parameters required are n, k, and p. The value of n is again the order of the graph. The value of k is one-sided neighbors, which can be estimated. The value of p is a rewiring probability, which is once again an estimation.

**Read before run**: This chunk generates a new model graph, but there is a stored one. Only use this if you wish to overwrite later. (check next chunk)
```{r}
n <- round(ma_pop_order * scale,0)
k <- round(ma_pop_size/ma_pop_order,0)
p <- 0.1443
set.seed(n)
g_pop_ws <- sample_smallworld(dim=1, size=n, nei=k, p=p)

g_sample_ws <- bfs_subgraph(g_pop_ws, scale = scale_sample_pop)
cat("Sample/pop scale: ", vcount(g_sample_ws)/vcount(g_pop_ws),"\n")
```

**Read before run**: If you want to use the previously generated graph, you can load it in this way.
```{r}
g_sample_ws <- read_graph(file.path(my_wd,"/model_graphs/swg_ma.csv"), "edgelist", directed=FALSE)
```

#### Analyze
The analysis function finds the information on the model network.

**Run**
```{r}
ws_sample_analysis <- graph_main_analysis(g_sample_ws)
```
**Run**
```{r}
cat("Order: ", ws_sample_analysis$order, "\n")
cat("Size: ", ws_sample_analysis$size, "\n")
cat("Degree range: ", ws_sample_analysis$degree_min, " - ", ws_sample_analysis$degree_max, "\n")
cat("Indegree range: ", ws_sample_analysis$indegree_min, " - ", ws_sample_analysis$indegree_max, "\n")
cat("Outdegree range: ", ws_sample_analysis$outdegree_min, " - ", ws_sample_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", ws_sample_analysis$clusters$weak, " / ", ws_sample_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", ws_sample_analysis$e_con, "\n")
cat("Vertex connectivity: ", ws_sample_analysis$v_con, "\n")
cat("Diameter: ", ws_sample_analysis$diameter, "\n")
cat("Average path length: ", ws_sample_analysis$apl, "\n")
cat("Transitivity: ", ws_sample_analysis$transitivity, "\n")
```

#### Save
**Read before run**: Watch out, don't unnecessarily overwrite the existing .csv graph representation.
```{r}
write_graph(g_sample_ws,file.path(my_wd,"/model_graphs/swg_ma.csv"),"edgelist")
```

### Barabási-Albert or preferential attachment graph analysis

#### Generate
The parameters required are n and m. The value of n is again the order of the graph. M is the degree of freedom, which can be estimated.

**Read before run**: This chunk generates a new model graph, but there is a stored one. Only use this if you wish to overwrite later. (check next chunk)
```{r}
n <- round(ma_pop_order * scale,0)
m <- round(ma_pop_size/ma_pop_order,0)
set.seed(n)
g_pop_pa <- sample_pa(n=n, m=m, directed=TRUE)

g_sample_pa <- bfs_subgraph(g_pop_pa, scale = scale_sample_pop)
cat("Sample/pop scale: ", vcount(g_sample_pa)/vcount(g_pop_pa),"\n")
```

**Read before run**: If you want to use the previously generated graph, you can load it in this way.
```{r}
g_sample_pa <- read_graph(file.path(my_wd,"/model_graphs/pa_ma.csv"), "edgelist", directed=TRUE)
```

#### Analyze
The analysis function finds the information on the model network.

**Run**
```{r}
pa_sample_analysis <- graph_main_analysis(g_sample_pa)
```
**Run**
```{r}
cat("Order: ", pa_sample_analysis$order, "\n")
cat("Size: ", pa_sample_analysis$size, "\n")
cat("Degree range: ", pa_sample_analysis$degree_min, " - ", pa_sample_analysis$degree_max, "\n")
cat("Indegree range: ", pa_sample_analysis$indegree_min, " - ", pa_sample_analysis$indegree_max, "\n")
cat("Outdegree range: ", pa_sample_analysis$outdegree_min, " - ", pa_sample_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", pa_sample_analysis$clusters$weak, " / ", pa_sample_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", pa_sample_analysis$e_con, "\n")
cat("Vertex connectivity: ", pa_sample_analysis$v_con, "\n")
cat("Diameter: ", pa_sample_analysis$diameter, "\n")
cat("Average path length: ", pa_sample_analysis$apl, "\n")
cat("Transitivity: ", pa_sample_analysis$transitivity, "\n")
```

#### Save
**Read before run**: Watch out, don't unnecessarily overwrite the existing .csv graph representation.
```{r}
write_graph(g_sample_pa,file.path(my_wd,"/model_graphs/pa_ts.csv"),"edgelist")
```

## Visualisations
Note: both layout and rendering calculations may take a long time

### Layout
The layouts are in separate cells, as execution takes quite long.

**Run**
```{r}
set.seed(length(V(g_sample_er))+length(V(g_sample_ws))+length(V(g_sample_pa)))
```

Note: Graphopt layouts produce odd, cross shaped plots, drl at least makes them interpretable. But please keep in mind that these layouts were not meant for such large graphs.

**Read before run**: This chunk generates a new graph layout, only use this if the order/size of the graph changed, as it takes quite long and previously generated layouts have been saved.
```{r}
#l_ma = layout_with_graphopt(g_ma, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_ma, file = file.path(my_wd,"layouts/l_ma.RDS"))
l_ma2 = layout_with_drl(g_ma)
saveRDS(l_ma2, file = file.path(my_wd,"layouts/l_ma2.RDS"))
```
**Read before run**: This chunk generates a new graph layout, only use this if the order/size of the graph changed, as it takes quite long and previously generated layouts have been saved.
```{r}
#l_er = layout_with_graphopt(g_sample_er, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_er, file = file.path(my_wd,"layouts/l_er_ma.RDS"))
l_er2 = layout_with_drl(g_sample_er)
saveRDS(l_er2, file = file.path(my_wd,"layouts/l_er_ma2.RDS"))
```
**Read before run**: This chunk generates a new graph layout, only use this if the order/size of the graph changed, as it takes quite long and previously generated layouts have been saved.
```{r}
#l_ws = layout_with_graphopt(g_sample_ws, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_ws, file = file.path(my_wd,"layouts/l_ws_ma.RDS"))
l_ws2 = layout_with_drl(g_sample_ws)
saveRDS(l_ws2, file = file.path(my_wd,"layouts/l_ws_ma2.RDS"))
```
**Read before run**: This chunk generates a new graph layout, only use this if the order/size of the graph changed, as it takes quite long and previously generated layouts have been saved.
```{r}
#l_pa = layout_with_graphopt(g_sample_pa, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_pa, file = file.path(my_wd,"layouts/l_pa_ma.RDS"))
l_pa2 = layout_with_drl(g_sample_pa)
saveRDS(l_pa2, file = file.path(my_wd,"layouts/l_pa_ma2.RDS"))
```

You can alternatively load them in from /layouts.
**Run**
```{r}
#l_ma <- readRDS(file.path(my_wd,"layouts/l_ma.RDS"))
#l_er <- readRDS(file.path(my_wd,"layouts/l_er_ma.RDS"))
#l_ws <- readRDS(file.path(my_wd,"layouts/l_ws_ma.RDS"))
#l_pa <- readRDS(file.path(my_wd,"layouts/l_pa_ma.RDS"))
l_ma2 <- readRDS(file.path(my_wd,"layouts/l_ma2.RDS"))
l_er2 <- readRDS(file.path(my_wd,"layouts/l_er_ma2.RDS"))
l_ws2 <- readRDS(file.path(my_wd,"layouts/l_ws_ma2.RDS"))
l_pa2 <- readRDS(file.path(my_wd,"layouts/l_pa_ma2.RDS"))
```

### Feature addition

Add the degrees
**Run**
```{r}
g_ma <- add_degree_attribute(g_ma)
g_sample_er <- add_degree_attribute(g_sample_er)
g_sample_ws <- add_degree_attribute(g_sample_ws)
g_sample_pa <- add_degree_attribute(g_sample_pa)
```

### Create graphs
This uses an external function for uniform representation.
**Run**
```{r}
g_graph_ma <- create_graph_visualization(g_ma, l_ma2, "purple")
g_graph_er <- create_graph_visualization(g_sample_er, l_er2, "blue")
g_graph_ws <- create_graph_visualization(g_sample_ws, l_ws2, "green")
g_graph_pa <- create_graph_visualization(g_sample_pa, l_pa2, "red")
```

**Read before run**: The rendering of the graphs takes quite long, if you want to take a look at them, check out the .png versions in /plots.
```{r}
g_graph_ma
ggsave(file.path(my_wd,"plots/ma.svg"), plot=g_graph_ma, device="svg", width=19, height=9)
```
**Read before run**: The rendering of the graphs takes quite long, if you want to take a look at them, check out the .png versions in /plots.
```{r}
g_graph_er
ggsave(file.path(my_wd,"plots/er_ma.svg"), plot=g_graph_er, device="svg", width=19, height=9)
```
**Read before run**: The rendering of the graphs takes quite long, if you want to take a look at them, check out the .png versions in /plots.
```{r}
g_graph_ws
ggsave(file.path(my_wd,"plots/ws_ma.svg"), plot=g_graph_ws, device="svg", width=19, height=9)
```
**Read before run**: The rendering of the graphs takes quite long, if you want to take a look at them, check out the .png versions in /plots.
```{r}
g_graph_pa
ggsave(file.path(my_wd,"plots/pa_ma.svg"), plot=g_graph_pa, device="svg", width=19, height=9)
```
