---
title: "Truth Social network analysis"
author: "Gábor T. Mozol"
date: "June, 2024"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Truth Social analysis

This notebook is responsible for the comparative analysis of the Truth Social data set acquired from the work of Gerard, Botzer and Weininger (2023).

The Truth Social graph is compared with model graphs:
* Erdős-Rényi or Gnp graph.
* Watts-Strogatz, a graph model with small-world properties.
* Barabási-Albert, a preferential attachment graph model.

The graph models are a 1:10 scale version of the estimated total size of the Truth Social network size in the Autumn of 2022. These model are then scraped using a similar BFS method as by Gerard et al. (2023). Based on our assumptions these samples are idealized reproductions of the Truth Social sample. The descriptive statistics of the data is copied over and the graph model descriptive statistics are calculated in this notebook.

## Libraries
The notebook uses the following libraries.
```{r warning=FALSE}
library(ggraph)
library(igraph)
library(svglite)
```
## Project setup
First, the wd needs to be set up.

Make sure to set you wd to the project folder.
```{r}
my_wd <- getwd()
project <- file.path(my_wd)
#usethis::create_project(project, open = FALSE)
usethis::proj_set(project)
source(file.path(my_wd,"functions/functions_analysis.R"))
```


## The Truth Social sample
The Truth Social sample relevant to graph topology analysis is stored and users.tsv and follows.tsv in the original dataset. Due to GitHub's size limitations this had to be split up and now they are contained in portions in folders of the respective name. See the attached readme file from the original authors to get a better understanding on the data. This split needs somen caretaking.

Then the data will be read into a graph and analysis function will return its descriptive statistics.

### Read and clean the data
```{r}
ts_data_path = paste(getwd(),"/data/truth_social_data", sep="")
ts_nodes <- read_tsv_files(file.path(ts_data_path, "users"))
ts_edges <- read_tsv_files(file.path(ts_data_path, "follows"))
```

```{r}
ts_nodes <- ts_nodes %>% select(id, username)
ts_edges <- ts_edges %>% select(follower, followed)
```

```{r}
colnames(ts_nodes)<-c("id", "username")
colnames(ts_edges)<-c("source", "target")
```

```{r}
ts_nodes <- check_append_missing_nodes(ts_edges, ts_nodes)
```

### Creat graph representation
```{r}
(g_ts <- igraph::graph_from_data_frame(
  d = ts_edges,
  directed = TRUE,
  vertices = ts_nodes
))
```
If you want to use the previously generated graph, you can load it in this way.
```{r}
g_ts <- read_graph(file.path(my_wd,"/graph/ts.csv"), "edgelist", directed=TRUE)
```

# Analysis
```{r}
g_ts_analysis <- graph_main_analysis(g_ts)
```

```{r}
cat("Order: ", g_ts_analysis$order, "\n")
cat("Size: ", g_ts_analysis$size, "\n")
cat("Degree range: ", g_ts_analysis$degree_min, " - ", g_ts_analysis$degree_max, "\n")
cat("Indegree range: ", g_ts_analysis$indegree_min, " - ", g_ts_analysis$indegree_max, "\n")
cat("Outdegree range: ", g_ts_analysis$outdegree_min, " - ", g_ts_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", g_ts_analysis$clusters$weak, " / ", g_ts_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", g_ts_analysis$e_con, "\n")
cat("Vertex connectivity: ", g_ts_analysis$v_con, "\n")
cat("Diameter: ", g_ts_analysis$diameter, "\n")
cat("Average path length: ", g_ts_analysis$apl, "\n")
cat("Transitivity: ", g_ts_analysis$transitivity, "\n")
```

### Save
```{r}
write_graph(g_ts,file.path(my_wd,"/graph/ts.csv"),"edgelist")
```


### Write Truth Social information
We store the Truth Social network information in the following variables.
```{r}
ts_sample_order <- g_ts_analysis$order
ts_sample_size <- g_ts_analysis$size
ts_sample_p <- 2*ts_sample_size/(ts_sample_order*(ts_sample_order-1))
```

The original paper scraped roughly 10.6% of the estimated user count. Giving us the population data
```{r}
scale_sample_pop <- 0.106
ts_pop_order <- round(ts_sample_order/scale_sample_pop,0)
ts_pop_size <- round(ts_sample_size/scale_sample_pop,0)
```

## Model graphs
Due to the large size of the original network we use a scaler for the order of the network.
```{r}
scale <- 1/10
```

### Erdős-Rényi or Gnp model graph analysis

#### Generate
The parameters required are n and p. The value of n is the order of the network, while p is the probability of a connection between any pair of nodes. This probability was calculated under the assumption that an average TS user might be connected to a 100 other users as follower or followee on average. We have the p value of the Truth Social network, ideally the generated network has a similar p value (i.e. within an order of magnitude).
```{r}
n <- round(ts_pop_order * scale,0)
p <- 0.000025 #(average 100 follower relationships in a 4.000.000 network)
set.seed(n)
g_pop_er <- sample_gnp(n, p, directed = TRUE)

g_sample_er <- bfs_subgraph(g_pop_er, scale = scale_sample_pop)
er_sample_order <- vcount(g_sample_er)
er_sample_size <- ecount(g_sample_er)
er_sample_p <- 2*er_sample_size/(er_sample_order*(er_sample_order-1))
cat("Sample/pop scale: ", vcount(g_sample_er)/vcount(g_pop_er),"\n")
cat("P value: ", er_sample_p,"\n")
cat("Target p value: ", ts_sample_p,"\n")
```
If you want to use the previously generated graph, you can load it in this way.
```{r}
g_sample_er <- read_graph(file.path(my_wd,"/model_graphs/gnp_ts.csv"), "edgelist", directed=TRUE)
```

#### Analyze
The analysis function finds the information on the model network.
```{r}
er_sample_analysis <- graph_main_analysis(g_sample_er)
```
```{r}
cat("Order: ", er_sample_analysis$order, "\n")
cat("Size: ", er_sample_analysis$size, "\n")
cat("Degree range: ", er_sample_analysis$degree_min, " - ", er_sample_analysis$degree_max, "\n")
cat("Indegree range: ", er_sample_analysis$indegree_min, " - ", er_sample_analysis$indegree_max, "\n")
cat("Outdegree range: ", er_sample_analysis$outdegree_min, " - ", er_sample_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", er_sample_analysis$clusters$weak, " / ", er_sample_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", er_sample_analysis$e_con, "\n")
cat("Vertex connectivity: ", er_sample_analysis$v_con, "\n")
cat("Diameter: ", er_sample_analysis$diameter, "\n")
cat("Average path length: ", er_sample_analysis$apl, "\n")
cat("Transitivity: ", er_sample_analysis$transitivity, "\n")
```

#### Save
```{r}
write_graph(g_sample_er,file.path(my_wd,"/model_graphs/gnp_ts.csv"),"edgelist")
```

### Watts-Strogatz or Small World Graph analysis
Disclaimer: The Watts-Strogatz network is NOT directed. So in and outdegree values cannot be compared by default.

#### Generate
The parameters required are n, k, and p. The value of n is again the order of the graph. The value of k is one-sided neighbors, which can be estimated. The value of p is a rewiring probability, which is once again an estimation.
```{r}
n <- round(ts_pop_order * scale,0)
k <- round(ts_pop_size/ts_pop_order,0)
p <- 0.1443
set.seed(n)
g_pop_ws <- sample_smallworld(dim=1, size=n, nei=k, p=p)

g_sample_ws <- bfs_subgraph(g_pop_ws, scale = scale_sample_pop)
cat("Sample/pop scale: ", vcount(g_sample_ws)/vcount(g_pop_ws),"\n")
```
If you want to use the previously generated graph, you can load it in this way.
```{r}
g_sample_ws <- read_graph(file.path(my_wd,"/model_graphs/swg_ts.csv"), "edgelist", directed=FALSE)
```

#### Analyze
The analysis function finds the information on the model network.
```{r}
ws_sample_analysis <- graph_main_analysis(g_sample_ws)
```
```{r}
cat("Order: ", ws_sample_analysis$order, "\n")
cat("Size: ", ws_sample_analysis$size, "\n")
cat("Degree range: ", ws_sample_analysis$degree_min, " - ", ws_sample_analysis$degree_max, "\n")
cat("Indegree range: ", ws_sample_analysis$indegree_min, " - ", ws_sample_analysis$indegree_max, "\n")
cat("Outdegree range: ", ws_sample_analysis$outdegree_min, " - ", ws_sample_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", ws_sample_analysis$clusters$weak, " / ", ws_sample_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", ws_sample_analysis$e_con, "\n")
cat("Vertex connectivity: ", ws_sample_analysis$v_con, "\n")
cat("Diameter: ", ws_sample_analysis$diameter, "\n")
cat("Average path length: ", ws_sample_analysis$apl, "\n")
cat("Transitivity: ", ws_sample_analysis$transitivity, "\n")
```

#### Save
```{r}
write_graph(g_sample_ws,file.path(my_wd,"/model_graphs/swg_ts.csv"),"edgelist")
```

### Barabási-Albert or preferential attachment graph analysis

#### Generate
The parameters required are n and m. The value of n is again the order of the graph. M is the degree of freedom, which can be estimated.
```{r}
n <- round(ts_pop_order * scale,0)
m <- round(ts_pop_size/ts_pop_order,0)
set.seed(n)
g_pop_pa <- sample_pa(n=n, m=m, directed=TRUE)

g_sample_pa <- bfs_subgraph(g_pop_pa, scale = scale_sample_pop)
cat("Sample/pop scale: ", vcount(g_sample_pa)/vcount(g_pop_pa),"\n")
```
If you want to use the previously generated graph, you can load it in this way.
```{r}
g_sample_pa <- read_graph(file.path(my_wd,"/model_graphs/pa_ts.csv"), "edgelist", directed=TRUE)
```

#### Analyze
The analysis function finds the information on the model network.
```{r}
pa_sample_analysis <- graph_main_analysis(g_sample_pa)
```
```{r}
cat("Order: ", pa_sample_analysis$order, "\n")
cat("Size: ", pa_sample_analysis$size, "\n")
cat("Degree range: ", pa_sample_analysis$degree_min, " - ", pa_sample_analysis$degree_max, "\n")
cat("Indegree range: ", pa_sample_analysis$indegree_min, " - ", pa_sample_analysis$indegree_max, "\n")
cat("Outdegree range: ", pa_sample_analysis$outdegree_min, " - ", pa_sample_analysis$outdegree_max, "\n")
cat("Clusters (weak / strong): ", pa_sample_analysis$clusters$weak, " / ", pa_sample_analysis$clusters$strong, "\n")
cat("Edge connectivity: ", pa_sample_analysis$e_con, "\n")
cat("Vertex connectivity: ", pa_sample_analysis$v_con, "\n")
cat("Diameter: ", pa_sample_analysis$diameter, "\n")
cat("Average path length: ", pa_sample_analysis$apl, "\n")
cat("Transitivity: ", pa_sample_analysis$transitivity, "\n")
```
#### Save
```{r}
write_graph(g_sample_pa,file.path(my_wd,"/model_graphs/pa_ts.csv"),"edgelist")
```

## Visualisations
Note: both layout and rendering calculations may take a long time

### Layout
The layouts are in separate cells, as execution takes quite long.
```{r}
set.seed(length(V(g_sample_er))+length(V(g_sample_ws))+length(V(g_sample_pa)))
```

```{r}
#l_er = layout_with_graphopt(g_sample_er, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_er, file = file.path(my_wd,"layouts/l_er_ts.RDS"))
l_er2 = layout_with_drl(g_sample_er)
saveRDS(l_er2, file = file.path(my_wd,"layouts/l_er_ts2.RDS"))
```
```{r}
#l_ws = layout_with_graphopt(g_sample_ws, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_ws, file = file.path(my_wd,"layouts/l_ws_ts.RDS"))
l_ws2 = layout_with_drl(g_sample_ws)
saveRDS(l_ws2, file = file.path(my_wd,"layouts/l_ws_ts2.RDS"))
```
```{r}
#l_pa = layout_with_graphopt(g_sample_pa, niter=100, charge = 0.05, mass = 30)
#saveRDS(l_pa, file = file.path(my_wd,"layouts/l_pa_ts.RDS"))
l_pa2 = layout_with_drl(g_sample_pa)
saveRDS(l_pa2, file = file.path(my_wd,"layouts/l_pa_ts2.RDS"))
```

You can alternatively load them in for /layouts.
```{r}
#l_er <- readRDS(file.path(my_wd,"layouts/l_er_ts.RDS"))
#l_ws <- readRDS(file.path(my_wd,"layouts/l_ws_ts.RDS"))
#l_pa <- readRDS(file.path(my_wd,"layouts/l_pa_ts.RDS"))
l_er2 <- readRDS(file.path(my_wd,"layouts/l_er_ts2.RDS"))
l_ws2 <- readRDS(file.path(my_wd,"layouts/l_ws_ts2.RDS"))
l_pa2 <- readRDS(file.path(my_wd,"layouts/l_pa_ts2.RDS"))
```

### Feature addition

Add the degrees
```{r}
g_sample_er <- add_degree_attribute(g_sample_er)
g_sample_ws <- add_degree_attribute(g_sample_ws)
g_sample_pa <- add_degree_attribute(g_sample_pa)
```

### Create graphs
This uses an external function for uniform representation.
```{r}
g_graph_er <- create_graph_visualization(g_sample_er, l_er2, "blue")
g_graph_ws <- create_graph_visualization(g_sample_ws, l_ws2, "green")
g_graph_pa <- create_graph_visualization(g_sample_pa, l_pa2, "red")
```

```{r}
g_graph_er
ggsave(file.path(my_wd,"plots/er_ts.svg"), plot=g_graph_er, device="svg", width=19, height=9)
```
```{r}
g_graph_ws
ggsave(file.path(my_wd,"plots/ws_ts.svg"), plot=g_graph_ws, device="svg", width=19, height=9)
```

```{r}
g_graph_pa
ggsave(file.path(my_wd,"plots/pa_ts.svg"), plot=g_graph_pa, device="svg", width=19, height=9)
```
